\documentclass[9pt]{article}
\usepackage[a4paper,
            top=1.5cm, bottom=1.5cm,
            left=1.5cm,  right=1.5cm]{geometry}

\usepackage{setspace}
\usepackage{microtype}
\usepackage[authoryear]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}

\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\captionsetup{font=small, labelfont=bf, justification=justified}

\usepackage[colorlinks=true,
            linkcolor=blue,
            citecolor=blue,
            urlcolor=blue]{hyperref}


\usepackage{footmisc}


\author{Maria Romagnoli and Lisa Chardon--Denizot}
\title{Applied labour economics}
\date{\today}
\begin{document}
\maketitle

\section{Introduction}
\label{sec:intro}


France maintains regional lists of shortage occupations (\textit{métiers en tension}),
which allow employers to hire non-European Economic Area (non-EEA) nationals without
the standard labour market test requiring proof that no domestic worker is available.
Since the 2024 immigration law \citep{ParFr2024}, residence permit renewals and
regularisation applications for non-EEA workers have been made contingent upon
employment in a listed occupation \citep{ServicePublic2025}. This creates a sharp
asymmetry: for a non-EEA worker, leaving a listed job entails the risk of losing
legal status, effectively collapsing their outside option to undocumented work or
the costly administrative process of obtaining a new permit. French nationals and
EEA citizens face no such constraint and retain the ability to switch employers or
occupations freely.

We ask whether firms exploit this asymmetry to set wages below competitive levels
for non-EEA workers in listed occupations, a monopsony effect induced by
the institutional design of the policy. Answering this question requires
disentangling three primary forces that simultaneously shape wages in shortage
occupations. First, labour scarcity may push firms to offer a shortage premium
to attract workers of any nationality. Second, taste-based discrimination may
independently depress non-EEA wages regardless of bargaining power. Third,
permit-conditioned immobility may generate a genuine monopsony markdown on top of
these two effects. Beyond these three channels, the policy also operates through
three additional mechanisms: a hiring cost reduction effect (the waiver of the
labour market test may allow employers to capture surplus from non-EEA workers, with
ambiguous wage consequences); a competition effect (listing may attract more
non-EEA workers into the occupation, increasing labour supply and depressing wages);
and a composition effect (listing may attract a different type of non-EEA
worker post-treatment, confounding the wage comparison). Separating the monopsony
channel from these five confounds is the central methodological challenge of
this paper.

We address it using a stacked difference-in-differences design with event studies,
comparing wage trajectories of non-EEA and EEA workers around listing events
relative to French nationals, on a culturally matched sample designed to minimise
contamination by a taste-based discrimination channel.
Our data cover the 2021, 2024, and 2025 iterations of the policy and are drawn
from the ForCE administrative database. We complement this main analysis with a
triple-differences regression for the 2021 reform, which provides a first estimate
of the monopsony effect in a simpler framework. Our main finding from the stacked
design is a precisely estimated null: we detect no statistically significant wage
markdown for non-EEA workers relative to their EEA counterparts following occupation
listing, for any reform cohort.

\section{Related Literature}
\label{sec:lit}

This paper sits at the intersection of three strands of literature: the economics of
employer monopsony, the labour market consequences of immigration policy, and the
measurement of wage discrimination.

\paragraph{Institutionally induced monopsony.}
A growing body of work shows that labour market institutions generate monopsony
power by restricting worker mobility. The most direct evidence comes from the Gulf
States, where \citet{Naidu2016} exploit a 2011 reform relaxing the UAE's
\textit{kafala} sponsorship system to show that, prior to reform, migrant workers
received only 51\% of their estimated marginal product; relaxing mobility constraints
raised this share to 72\%. The mechanism (legal status tied to a single employer)
is structurally analogous to the French shortage-list regime, although the
French constraint is partial rather than absolute. \citet{Amior2024} provide
complementary evidence on German administrative data, showing that immigration
strengthens firm wage-setting power at the bottom of the pay distribution. Our
contribution is to test whether an occupation-level mobility restriction of moderate
intensity produces similar effects in a high-regulation European labour market.

\paragraph{Selective immigration policy and native--migrant wage gaps.}
The paper most closely related to ours is \citet{Signorelli2024}, which evaluates
the first iteration of the French shortage list policy introduced in 2008. Using a
difference-in-differences strategy, she finds that the policy-induced influx of
migrant workers depressed migrant wages more than native wages, a differential
consistent with our monopsony hypothesis. We extend her analysis to the 2021--2025
period, exploit three distinct reform cohorts, and bring an explicit conceptual
framework that separates the monopsony channel from the shortage premium and
discrimination channels. More broadly, our work relates to studies of how immigration
policy shapes labour market outcomes \citep{Peri2012, Dustmann2017}, with a specific
focus on wage consequences for migrant workers themselves rather than for native
incumbents.

\paragraph{Discrimination and wage penalties.}
Taste-based discrimination \citep{Becker57} can generate wage penalties that mimic
monopsony markdowns but operate through entirely different mechanisms. \citet{Glover2017}
provide quasi-experimental evidence from French grocery stores that manager bias against
minority workers produces measurable productivity and wage gaps. Accounting for this
channel is essential for causal interpretation. Our culturally matched sample strategy
(using EEA workers who share broad cultural backgrounds with their non-EEA
counterparts but face no permit constraint as an auxiliary control group) is
designed precisely to absorb time-invariant discrimination. Time-varying discrimination
is further addressed by the sample construction, which pairs nationalities on
ethno-linguistic proximity.


\section{Data}
\label{sec:data}


\subsection{Administrative data: ForCE}

Our primary data source is ForCE (\textit{Formation, Chômage et Emploi}),
a French administrative database managed by France Travail that links unemployment
spells, job contracts, and individual characteristics for all workers registered with
the public employment service. We use the 2025S1 vintage and focus on the period
2019--2025, which provides at least one pre-reform year before each of the three
listing events.

We draw on two tables within ForCE. The \textit{Fichier Historique (FH)} provides
worker-level socio-demographic characteristics recorded at the unemployment-spell
level, including nationality (\texttt{NATION}), sex, date of birth, diploma level,
qualification, and cumulative experience in the target occupation (\texttt{ROME} code).
When a characteristic varies across spells, we retain the most recent observed value.
The Mouvement de Main-d'{\OE}uvre (MMO) records contract-level information,
including contract start and end dates, establishment identifier, full- or part-time
status, base monthly salary, and base salary for a full month of work. We restrict
MMO to contracts longer than 30 days beginning after January 2019. Department codes
in MMO are used to recover the region of each contract. The two tables are merged on
the \texttt{force\_id} unique identifier.

Because a worker may hold multiple contracts over the observation window, the final
dataset is structured at the contract level. Worker characteristics from FH
are attached to each contract via the ForCE key, so workers with multiple contracts
appear repeatedly. Specifications weighted at the worker level are reported as a
robustness check. The final sample covers 60,214,080 contracts for French workers,
681,184 for EEA workers, and 884,757 for non-EEA workers (Table %add balance checks).

\subsection{Descriptive evidence}

Table balance checks presents balance statistics by treatment and permit status.
Several patterns are noteworthy. First, workers in listed occupations are on average
older, earn higher wages (€1,771 vs.\ €1,588 per month), and are more likely to be
male (61\% vs.\ 49\%), consistent with listed occupations being concentrated in
skilled trades and technical services. Second, non-EEA workers earn less on average
than both French nationals and EEA workers (€1,560 vs.\ €1,720 and €1,653,
respectively), but report more experience (5.59 years vs.\ 4.30 and 4.08). This
experience premium without a commensurate wage premium is suggestive of a wage gap
that cannot be explained by observable human capital alone, and motivates the more
careful causal analysis that follows. Third, EEA and non-EEA samples are
substantially smaller than the French reference group, which limits statistical power
for the culturally matched subsamples.

Figure %add boxplot
displays the distribution of monthly wages by nationality
group and listing status. Wages in listed occupations are higher for all groups,
confirming the existence of a shortage premium. The wage gap between listed and
unlisted occupations is visually similar across groups, suggesting that the shortage
premium is not substantially heterogeneous by permit status at the raw data level.
Recovering the monopsony component net of this premium requires the causal designs
described below.

\subsection{Nationality classification and culturally matched samples}

Non-EEA workers are identified through the \texttt{NATION} codes in FH. Raw
non-EEA versus EEA comparisons conflate the permit-constraint channel with cultural,
religious, or human-capital differences that may independently affect wages. We
restrict the main estimation sample to culturally matched pairs of non-EEA
and EEA nationalities that share broad ethno-linguistic or historical backgrounds
but differ in permit status. Table nationalities in the Appendix details four
matched pairs: (i) former Yugoslav nationalities (Montenegro, Serbia, Bosnia and
Herzegovina, Kosovo, North Macedonia) matched with Croatian and Slovenian nationals;
(ii) the same Yugoslav group matched with Bulgarian and Romanian nationals;
(iii) Turkish nationals matched with Greek, Bulgarian, Romanian, and Cypriot nationals
(with a residual religious bias caveat); and (iv) residual European non-EEA nationals
matched with Baltic and Central European EEA nationals. French nationals serve as
the reference group throughout.

\subsection{Listing status}

Listing status at the occupation-region-year level is constructed from the
\textit{arrêtés} published in the \textit{Journal Officiel de la République
Française} for the 2021, 2024, and 2025 reforms. Shortage lists are defined at
the \textit{famille professionnelle} (FAP) code level in the policy texts; we map
FAP codes to the PCS codes recorded in MMO to assign listing status to each
contract. Table listing summarises the number of listed occupations by
region and reform year. Substantial regional variation exists: Île-de-France went
from 30 listed occupations in 2021 to 17 in 2025, while PACA moved from 26 to 30.
This variation, across occupations, regions, and time, is the identifying variation
exploited by our empirical designs.

\section{Model}
\label{sec:model}

\subsection{Conceptual framework}

We model wages as the outcome of Nash bargaining between firms and workers.
The negotiated wage is a weighted average of the firm's outside option (zero profit)
and the worker's outside option, with the firm's bargaining weight increasing in the
tightness of labour supply. For French nationals and EEA workers, the outside option
approximates the competitive market wage, since both groups can freely switch
employers or occupations. For non-EEA workers in listed occupations, leaving the
job entails forfeiting the residence permit, so the effective outside option is
undocumented work or the risk-weighted cost of obtaining a new permit. Under this framework, listing an occupation generates
a divergence in bargaining positions: the firm can extract more surplus from non-EEA
workers than from EEA workers doing the same job, producing a wage markdown.
Whether this markdown is empirically detectable depends on the magnitude of the
bargaining power differential and the countervailing effects described above.

Table~\ref{tab:mechanisms} summarises all six mechanisms through which listing may
affect the wages of non-EEA workers relative to other groups, and how each is
handled by our identification strategy.



\begin{table}[H]

    \centering
\caption{Mechanisms, directions, and identification.}
\label{tab:mechanisms}
\small
\begin{tabular}{p{2.5cm} p{5.5cm} p{1.2cm} p{4cm}}
\toprule
Effect & Mechanism & Direction & Status in our design \\
\midrule
Monopsony       & Listing collapses non-EEA outside option; EEA workers unaffected
                & $\downarrow$ & Target parameter \\[4pt]
Competition     & More non-EEA workers enter listed occupations post-listing,
                  increasing labour supply
                & $\downarrow$ & Confounded \\[4pt]
Hiring cost     & Labour market test waived for non-EEA only; employer captures
                  surplus
                & ambiguous & Confounded \\[4pt]
Composition     & Listing attracts a different type of non-EEA worker post-treatment
                & $\downarrow$ & Confounded \\[4pt]
Shortage        & Labour scarcity raises wages for all workers
                & $\uparrow$ & Differenced out by EEA coefficient \\[4pt]
Discrimination  & Taste-based penalties depress non-EEA wages;
                  if time-invariant, absorbed by FE;
                  if time-varying, managed by matched sample
                & $\downarrow$ & Largely controlled \\
\bottomrule
\end{tabular}
\begin{minipage}{\linewidth}
\vspace{4pt}
\footnotesize\textit{Notes:} $\downarrow$ indicates downward pressure on non-EEA wages
relative to other groups; $\uparrow$ indicates upward pressure for all groups.
\end{minipage}
\end{table}


\subsection{Specification A: triple differences}

As a first pass, we estimate a triple-differences (DDD) regression for the 2021
reform. The specification compares wage changes across three dimensions: nationality
group, treatment status (occupation-region cells listed after April 2021 versus
those that were not), and the post-2021 period. Formally:

\begin{align}
\ln w_{ijrt} = \;& \alpha + \beta_1 \mathrm{EEA}_i + \beta_2 \mathrm{NonEEA}_i
  + \beta_3 \mathrm{post}_t + \mathbf{X}'_{ijrt}\gamma \notag \\
  &+ \beta_4 (\mathrm{French}_i \times \mathrm{treatment}_{jr} \times \mathrm{post}_t)
   + \beta_5 (\mathrm{EEA}_i \times \mathrm{treatment}_{jr} \times \mathrm{post}_t) \notag \\
  &+ \beta_6 (\mathrm{NonEEA}_i \times \mathrm{treatment}_{jr} \times \mathrm{post}_t)
   + \delta_{jr} + \delta_t + \varepsilon_{ijrt},
\label{eq:ddd}
\end{align}

\noindent where the monopsony parameter of interest is $\beta_6$: the differential
wage change for non-EEA workers in newly-listed cells after April 2021, relative to
French nationals in the same cells. The coefficient $\beta_5$ provides a placebo-style
check: under the hypothesis that only permit-constrained workers are affected,
$\beta_5$ should be zero. Individual controls $\mathbf{X}_{ijrt}$ include sex, diploma,
and experience in months. Fixed effects $\delta_{jr}$ absorb time-invariant
occupation-region heterogeneity; $\delta_t$ absorbs common year effects. Standard
errors are clustered at the occupation-by-region level.

\subsection{Specification B: stacked event-study}

The DDD estimate may be contaminated by pre-existing differential trends and by
the use of late-listed occupation-region cells as implicit controls for earlier
cohorts, a well-documented source of bias in staggered DiD settings
\citep{Callaway2021, Sun2021}.
Therefore, ou preferred empirical strategy builds on a stacked difference-in-differences design,
following the approach advocated by \citet{Cengiz2019} and more recently
formalized for multi-period settings by \citet{Baker2022}.\footnote{%
  The stacked design constructs a separate clean "mini-panel" for each
  listing event (2021, 2024, and 2025) defined by a window
  around the relevant reform, and then pools these sub-experiments.
  This avoids the contamination of late-treated units serving as controls
  for early-treated units that plagues two-way fixed-effects estimators
  in heterogeneous-treatment settings \citep{Callaway2021, Sun2021}.}
For practical estimation,we estimate two separate difference-in-differences regressions
and recover the quantity of interest by comparing the resulting coefficients using the linearity of
conditional expectation.

For each listing cohort $c \in \{2021, 2024, 2025\}$, we estimate the
following stacked event-study regression:

\begin{equation}
  \ln w_{ijrts} =
    \sum_{g \neq -1} \beta_k \bigl(\mathbf{1}[\text{group}_i = g]
      \times \mathbf{1}[\text{policy}_{jrt} = k]\bigr)
    + \mathbf{X}_{ijrt}'\,\gamma
    + \delta_{jrt}
    + \varepsilon_{ijrts},
  \label{eq:eventstudy}
\end{equation}

\noindent
where $w_{ijrts}$ is the monthly base wage of worker $i$ in occupation $j$,
region $r$, at time $t$, belonging to stacked sub-experiment $s$.
The variable $\text{policy}_{jrt} = k$ indexes event time relative to the first
listing of occupation $j$ in region $r$, normalized so that $k = -1$ is the
omitted group (French). $\text{group}_i \in
\{\text{non-EEA},\, \text{EEA},\, \text{French}\}$ identifies the
nationality group of the worker, with French nationals serving as the
reference category.

The individual control vector $\mathbf{X}_{ijrt}$ includes sex, diploma
level, and cumulative experience in months. The fixed-effect structure
$\delta_{jrt}$ absorbs occupation-by-region-by-year variation, which is
the most demanding specification able to control for time-varying labour
scarcity at the occupation-region level, i.e. the shortage premium
effect. Because this high-dimensional fixed
effect is mechanically collinear with the listing indicator, we repeat the
analysis with coarser fixed-effect sets as robustness checks.

The regression is estimated separately for each listing cohort using the
\texttt{feols} function from the \texttt{fixest} package in \textsf{R}
\citep{Berge2018}. Standard errors are clustered at the
occupation-by-region level to account for within-cluster correlation in
both treatment assignment and wages.

\subsection{Parameter of interest and identification}

Let $\hat\beta_k^{\text{non-EEA}}$ and $\hat\beta_k^{\text{EEA}}$ denote
the estimated event-study coefficients for non-EEA and EEA workers
respectively, each relative to French nationals. The parameter of interest
at event time $k$ is the differential post-listing wage effect:

\begin{equation}
  \Delta_k =
  \hat\beta_k^{\text{non-EEA}} - \hat\beta_k^{\text{EEA}}.
  \label{eq:delta}
\end{equation}

Under our hypothesis, we expect $\Delta_k < 0$ for $k \geq 0$:
non-EEA workers should experience a relative wage markdown upon listing,
once the EEA coefficient absorbs any shortage premium accruing to both
groups. The statistical significance of $\Delta_k$ is assessed via the
delta method, which yields an asymptotically valid variance estimate for
linear combinations of estimated coefficients.

Identification rests on four assumptions.
The primary concern is parallel pre-trends: absent listing, non-EEA and
EEA wage trajectories should be parallel to those of French nationals
within the same occupation-region cell. This assumption is testable and
forms the main diagnostic reported below. A second threat is endogenous
worker selection into listed occupations following the reform; to the
extent that selection is positive on unobservables, any estimated markdown
is a lower bound on the true monopsony effect. Third, we assume no general
equilibrium spillovers from listed to unlisted occupations at the
occupation-region level. Fourth, we treat listing events as locally
exogenous, conditional on fixed effects.

\section{Results}
\label{sec:results}


\subsection{Triple-differences estimates}

Table results reports the DDD estimates for the 2021 reform. The baseline
wage gaps are substantial: EEA workers earn approximately 6.3\% more than French
nationals conditional on controls, while non-EEA workers earn 4.6\% less
($\hat\beta_1 = 0.063$, $\hat\beta_2 = -0.046$, both significant at the 0.1\%
level). The listing effect for French workers ($\hat\beta_4$) is negative and
significant ($-0.026$, $p = 0.002$), suggesting that French workers in newly-listed
cells experienced a relative wage decline after 2021, possibly reflecting general
equilibrium effects of increased labour supply in those occupations.

The EEA placebo coefficient $\hat\beta_5$ is small and insignificant
($-0.005$, $p = 0.837$), consistent with the hypothesis that EEA workers, who
are unaffected by the permit constraint, did not experience a differential wage
change upon listing. The non-EEA coefficient $\hat\beta_6$ is negative and
significant at the 5\% level ($-0.041$, $p = 0.048$), with a magnitude suggesting
a 4.1\% wage markdown relative to French workers in the same occupation-region
following listing. This estimate is broadly consistent with the monopsony hypothesis.

However, this result must be interpreted with caution. The pre-trend analysis
reveals that treated and control
cells were already on diverging wage trajectories before the reform, particularly for
EEA and non-EEA subgroups, which exhibit high month-to-month volatility attributable
to small sample sizes. This pre-existing divergence implies that the parallel trends
assumption underlying the DDD is likely violated, and the $\hat\beta_6$ estimate
may partially capture a pre-existing gap rather than a causal listing effect.

%% Add results 24 and 25
\subsection{Event-study estimates}

Figure %Include event studies graphs%
display the event-study
graphs for the 2021, 2024, and 2025 listing cohorts. Each figure presents
three panels: the estimated $\hat\beta_k^{\text{non-EEA}}$ series (non-EEA
vs.\ French), the estimated $\hat\beta_k^{\text{EEA}}$ series (EEA
vs.\ French), and the implied differential $\Delta_k$ computed via the
delta method. Ninety-five-percent confidence intervals are reported
throughout.

For all three cohorts, the pre-treatment coefficients are statistically
indistinguishable from zero and display no discernible trend, supporting
the parallel pre-trends assumption. This lends credibility to the
identification strategy, even though formal pre-trend tests cannot be
interpreted as proof of the counterfactual assumption.

In the post-treatment periods, neither $\hat\beta_k^{\text{non-EEA}}$ nor
$\hat\beta_k^{\text{EEA}}$ departs significantly from zero for any cohort.
Point estimates are small in magnitude and confidence intervals are wide,
which may reflect genuine heterogeneity in
treatment effects across regions and occupations.

\subsection{Differential effects and delta-method tests}

Table %Include results table
reports the post-listing average of $\Delta_k$ for
each cohort, together with the corresponding standard error and $p$-value
obtained from the delta method.

For the 2021 cohort, no statistically significant differential is detected.
Point estimates of $\Delta_k$ fluctuate around zero in the post-period,
and the null hypothesis $H_0: \Delta_k = 0$ cannot be rejected at any
conventional significance level.  Results for the 2024 cohort are qualitatively similar, with no significant
differential wage effect for non-EEA relative to EEA workers following
listing.

The 2025 cohort yields a weakly significant result at the ten-percent
level for certain post-listing periods. Interestingly, the sign of
$\Delta_k$ is \emph{positive} in this cohort: the point estimate suggests
a slightly higher relative wage for non-EEA workers following listing,
compared to their EEA counterparts. This finding is, if anything,
inconsistent with the monopsony hypothesis; it is more plausibly
attributable to sampling noise given the short post-treatment window
available in the most recent vintage of ForCE.

\subsection{Robustness}

Coarser fixed-effect specifications (e.g., replacing occupation-by-region-by-year
with occupation-by-year fixed effects) do not materially alter the conclusions.

\section{Conclusion}
\label{sec:conclusion}

This paper investigates whether French shortage-occupation lists create
rents that allow firms to depress the wages of non-EEA workers,
whose legal status is tied to remaining in a listed occupation. We embed
this question in a Nash bargaining framework predicting a wage differential
between non-EEA and EEA or French workers following listing, and test it
using a stacked event-study design on ForCE administrative data covering
the 2021, 2024, and 2025 iterations of the policy.

Our main finding is a null result: we do not detect a statistically
significant wage markdown for non-EEA workers relative to EEA workers
following occupation listing, for any of the three reform cohorts
examined. Pre-treatment parallel trends hold throughout, suggesting that
the empirical design is internally valid. The absence of a significant
post-treatment differential is therefore not a mechanical artefact of
the estimation strategy.

Several interpretations of this null are plausible. First, firms may
not exploit workers' permit-conditioned immobility in the short run,
either because wage-setting in these occupations is governed by
collective agreements that limit bilateral bargaining, or because
reputational concerns discourage overt wage discrimination. Second,
non-EEA workers in shortage occupations may self-select on unobservable
productivity that offsets the monopsony markdown, generating an upward
bias that masks a negative structural effect. Third, the window available
for the most recent cohorts (in particular 2025) is too short to
detect dynamic wage effects that may materialise over multiple years.

Our results speak to a growing literature on institutionally induced
monopsony \citep{Naidu2016, Amior2024}. Unlike the UAE kafala setting
studied by \citet{Naidu2016}, the French shortage-list mechanism does not
prohibit job mobility outright; it merely raises its costs. Our null
result is consistent with the hypothesis that moderate rather than
absolute mobility restrictions are insufficient to generate detectable
wage markdowns in a high-regulation labour market where sectoral
bargaining compresses wage dispersion.

Future work should extend the observation window, particularly for the
2024 and 2025 cohorts, as longer panels will improve power and allow
examination of the dynamic path of wage effects. Complementing the wage
analysis with evidence on employment margins (whether listing affects
non-EEA hiring rates or contract duration differentially) would provide
a richer picture of how firms respond to the policy. Finally, exploiting
variation in the stringency of local enforcement or in the share of
workers covered by collective agreements could help disentangle the
mechanisms underlying the null wage result.

    \bibliography{biblio}
    \bibliographystyle{apalike}
\end{document}